---
title: "Bee Respiration for Susie and Stacey"
author: "Callin Switzer"
date: "11/4/2016, Update: 12/13/2016, Update: 28 Jan 2017"
output: html_document
---


---

#### Update 28 Jan 2017

#### Summary of findings:
1. % load and total mass are too collinear to include in the same model.
2. Treatment and % load are too collinear to include in the same model.
3. For the remainder of the modeling, I'll use only % load


#### Differences between model with Treatment vs. model with % load
- Treatment order is not a significant predictor of wingbeat frequency in the model with percent load (though it is very close, p ~ 0.07)

- percent load and treatment suggest very similar models in terms of wingbeat frequency

- Treatment order is not a strong predictor of stroke amplitude in the model with percent load





wingbeat freq = size
Met rate  = mass carrying
amplitude = percent loading


Susie’s (hypotheses):
lm( frequency ~ ITspan  +  trt order)

lm( metabolism ~ total mass)

lm( stroke amplitude ~ %load  + trt) (those are probably the same thing)


* Redo the model with %load and/or total mass (probably the best predictor of metabolism)
* Check VIF
* Think about how to make figures
* How heavy the bee is loaded overall, affects how much the amplitude and frequency change

- Questions to look into:
** A really heavy loaded bee will have a smaller change in metabolism vs. light bee
** Suggests a non-linear relationship or random slopes that depend on bee size — the cost for additional weight goes down as you get heavier loaded. — not the case for the change in amplitude  (change in load vs. change in amplitude is linear) — not the case for metabolism or wingbeat frequency.

** Hypotheses: 
1. Wingbeat freq moves around for a number of reasons (tired, cold, added mass)
2. Stroke amplitude rescues wingbeat frequency
3. As you get really close to your max, adding more weight doesn’t seem to cost as much.




---

Analyses of respirometry data:

1.	How does carrying a load during flight affect bumblebees’?
     a.	How does load affect respiratory rate?
     b.	How does load affect wingbeat frequency?
     c.	How does load affect stroke amplitude?
     d.	How does load affect wing velocity?

2.	Which flight measurements are most closely associated with respiratory rate in bees – wingbeat frequency, stroke amplitude, or wing velocity?

---

### Setup
```{r setup, echo=FALSE}
knitr::opts_knit$set(root.dir = "~/Dropbox/dataAnalysisForOthers/SusieBeeWingAndResp/BeeResp_Oct2016/")
```

Install required packages and read in data
Define custom function for evaluating VIF with multilevel models

```{r, message=FALSE, warning = FALSE}
ipak <- function(pkg){
     new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
     if(length(new.pkg)) install.packages(new.pkg, dependencies = TRUE)
     sapply(pkg, require, character.only = TRUE)
}

packages <- c('ggplot2', 'car', 'lme4', 'gsheet', "MASS", 'influence.ME')
ipak(packages)

theme_set(theme_bw())

# read in data -- google sheet called "Bumble mumble grumble"
bdta <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1GUUoFq41Ep3sJNFXiyMcp7fZhYmQCzTcLnhVXdr4WRo/edit?usp=sharing")
```


#### Function for calculating vif for lmer
(used later)
```{r}
vif.mer <- function (fit) {
    ## adapted from rms::vif
    v <- vcov(fit)
    nam <- names(fixef(fit))

    ## exclude intercepts
    ns <- sum(1 * (nam == "Intercept" | nam == "(Intercept)"))
    if (ns > 0) {
        v <- v[-(1:ns), -(1:ns), drop = FALSE]
        nam <- nam[-(1:ns)]
    }
    
    d <- diag(v)^0.5
    v <- diag(solve(v/(d %o% d)))
    names(v) <- nam
    v
}

```


## Data Overview

```{r}

# calculate percent load
# calculated as total mass / starved mass
bdta$percLoad <- with(bdta, (Mstarved + load) / Mstarved)

# center percent load
bdta$percLoad <- scale(bdta$percLoad, center = TRUE, scale = FALSE)
hist(bdta$percLoad)

# calculate total mass
bdta$totalMass <- with(bdta, Mstarved + load)

summary(bdta)

ggplot(bdta, aes(x = percLoad, fill = Treatment)) + 
     geom_histogram()  
# seems like percent load and trt are very related
```


```{r}
# visualize % load vs total mass
ggplot(bdta, aes(x = percLoad, y = Mstarved + load)) + 
     geom_point()
# seem to be strongly associated
```

```{r}
# see number of observations per bee
table(bdta$Bee.ID) # each bee has two observations

# get number of unque bees
length(table(bdta$Bee.ID))

# get treatment orders
trtOrders_loadedSecond <- sapply(X = unique(bdta$Bee.ID), FUN = function(x){
     tmp = bdta[bdta$Bee.ID == x, c("Treatment.order", "Treatment")]
     if("2_L" %in% paste(tmp[,1], tmp[,2], sep = "_")){
          loadedSecond = TRUE
     }
     else loadedSecond = FALSE
     return(loadedSecond)
})

table(trtOrders_loadedSecond)



# visualize scatterplot
car::scatterplotMatrix(bdta[, 4:13])
```

### Variance Inflation Factor and PCA

```{r}
# VIF  is high among the bee size predictors
# note: this function is almost the same as car::vif
vif.mer(lmer(av.resp..CO2.mL.hr. ~ Treatment + Treatment.order + Mstarved + IT.Span + totalMass +  single.wing.area + percLoad +  (1|Bee.ID), data = bdta))

car::scatterplotMatrix(bdta[, c("Mstarved", "IT.Span", "single.wing.area")])

# principle components
aa = prcomp(bdta[, c("Mstarved", "IT.Span", "single.wing.area")], center = TRUE, scale = TRUE)
summary(aa) # 1st pc explains ~95% of the variance in the three measurements of size

biplot(aa) # shows that all three size measurement are correlated

# note, I changed the signs of the predictions so that higher PC1 values 
# correspond to bigger bees
p1 = -predict(aa)[,1] 

# add PC1 scores to dataset
bdta$size_pca1 = p1

# show scatterplot matrix to see correlations among size predictors
car::scatterplotMatrix(bdta[, c("Mstarved", "IT.Span", "single.wing.area", "size_pca1", "percLoad")])

# check VIF one more time
# VIF  is high among the bee size predictors
vif.mer(lmer(av.resp..CO2.mL.hr. ~ Treatment + Treatment.order + size_pca1 + percLoad + totalMass +  (1|Bee.ID), data = bdta))


# remove treatment 
vif.mer(lmer(av.resp..CO2.mL.hr. ~ Treatment.order + size_pca1 + percLoad + totalMass + (1|Bee.ID), data = bdta))
# looks like we cannot have total mass and percentLoad in the same model, according to VIF

# remove total mass 
vif.mer(lmer(av.resp..CO2.mL.hr. ~ Treatment.order + size_pca1 + percLoad +  (1|Bee.ID), data = bdta))
```



---
Part I: How does carrying a load during flight affect bumblebees’?
a.	How does %load affect respiratory rate?
b.	How does %load affect wingbeat frequency?
c.	How does %load affect stroke amplitude?
d.	How does %load affect wing velocity?
In the models, we'll account for differences in bee size and the order of the treatment
---

## Multilevel models

```{r}
# a few house-keeping issues to make models easier to read

# convert trt order to factor
bdta$Treatment.order <- as.factor(as.character(bdta$Treatment.order))

# change reference level to unloaded
bdta$Treatment <- factor(bdta$Treatment, levels = c("UL", "L"))
```

## Resp Rate

```{r, cache = TRUE}
# make a full model with all two-way interactions
m1 <- lmer(av.resp..CO2.mL.hr. ~  (size_pca1 +  Treatment.order + percLoad)^2 + (1|Bee.ID), data = bdta)

summary(m1)


### LRT's for interactions
m2.0 <- update(m1, .~. - Treatment.order:percLoad)

anova(m1, m2.0) # likelihood ratio test for interaction of treatment order
summary(m2.0)


m2.1 <- update(m2.0,.~. - size_pca1:Treatment.order)
anova(m2.0, m2.1)
summary(m2.1)


m2.2 <- update(m2.1, .~. - size_pca1:percLoad)
anova(m2.1, m2.2) ## drop all interactions

summary(m2.2)


# renaming model to simplify later typing
m2 <- m2.2

##### LRTs for main effects
## Treatment Order
m3 <- update(m2, .~. - Treatment.order)
anova(m2, m3, test = "Chi") # drop trt order (different than model with treatment)

# LRT for size
m4 <- update(m3, .~. - size_pca1)
anova(m3, m4) # keep size

# LRT for Treatment (load)
m5 <- update(m3, .~. - percLoad)
anova(m3, m5) # keep percent load

# summarize final model for paper
summary(m3)



# write output
summary(m3)$coefficients  
write.csv( summary(m3)$coefficients, file = "RespCoefs_percLoad.csv" )


# diagnostics
# qq plot
qqnorm(resid(m3), main = "")
qqline(resid(m3)) # good

# residual plot
plot(fitted(m3), resid(m3), xlab = "fitted", ylab = "residuals")
abline(0,0)

# check cook's distance -- should be less than 1, so we're good
infl <- influence(m3, obs = TRUE)
plot(infl, which = 'cook')

# QQPlot for group-level effects
qqnorm(ranef(m3)$Bee.ID[[1]], main="Normal Q-Q plot for random effects")
qqline(ranef(m3)$Bee.ID[[1]]) # looks good
```

## Wingbeat Freq

```{r, cache = TRUE}
# make a full model with all two-way interactions
m1 <- lmer(wbf.aud. ~  (size_pca1 +  Treatment.order + percLoad)^2 + (1|Bee.ID), data = bdta)

summary(m1)


### LRT's for interactions
m2.0 <- update(m1, .~. - size_pca1:percLoad )

anova(m1, m2.0)
summary(m2.0)


m2.1 <- update(m2.0,.~. - size_pca1:Treatment.order)
anova(m2.0, m2.1)
summary(m2.1)


m2.2 <- update(m2.1, .~. - Treatment.order:percLoad)
anova(m2.1, m2.2) ## drop all interactions

summary(m2.2)


# renaming model to simplify later typing
m2 <- m2.2

##### LRTs for main effects
## Treatment Order
m3 <- update(m2, .~. - Treatment.order)
anova(m2, m3, test = "Chi") # keep treatment order

# LRT for size
m4 <- update(m2, .~. - size_pca1)
anova(m2, m4) # keep size_pca1

# LRT for Treatment (load)
m5 <- update(m2, .~. - percLoad)
anova(m2, m5) # keep percLoad

# summarize final model for paper
summary(m2)

# write output
summary(m2)$coefficients  
write.csv( summary(m2)$coefficients, file = "FreqCoefs_percLoad.csv" )


# diagnostics
# qq plot
qqnorm(resid(m2), main = "")
qqline(resid(m2)) # good

# check influence
plot(influence(m2, obs = TRUE ), which = 'cook')

# residual plot
plot(fitted(m2), residuals(m2, "deviance"), xlab = "fitted", ylab = "residuals")
abline(0,0)




# QQPlot for group-level effects
qqnorm(ranef(m2)$Bee.ID[[1]], main="Normal Q-Q plot for random effects")
qqline(ranef(m2)$Bee.ID[[1]]) # looks good
```


## Stroke Amplitude
```{r}
# make a full model with all two-way interactions
m1 <- lmer(stroke.amplitude ~  (size_pca1 +  Treatment.order + percLoad)^2 + (1|Bee.ID), data = bdta)

summary(m1)


### LRT's for interactions
m2.0 <- update(m1, .~. - Treatment.order:percLoad )

anova(m1, m2.0)
summary(m2.0)


m2.1 <- update(m2.0,.~. - size_pca1:Treatment.order)
anova(m2.0, m2.1)
summary(m2.1)


m2.2 <- update(m2.1, .~. - size_pca1:percLoad)
anova(m2.1, m2.2) ## drop all interactions

summary(m2.2)


# renaming model to simplify later typing
m2 <- m2.2

##### LRTs for main effects
## size
m3 <- update(m2, .~. - size_pca1)
anova(m2, m3, test = "Chi") # keep size_pca1
summary(m3)

# trt order
m4 <- update(m2, .~. - Treatment.order)
anova(m2, m4) # drop treatment order

# LRT for Treatment (load)
m5 <- update(m4, .~. - percLoad)
anova(m4, m5)

# summarize final model for paper
summary(m4)

# write output
summary(m4)$coefficients  
write.csv( summary(m4)$coefficients, file = "AmpCoefs_percLoad.csv" )


# diagnostics
# qq plot
qqnorm(resid(m4), main = "")
qqline(resid(m4)) # ok

# residual plot
plot(fitted(m4), resid(m3), xlab = "fitted", ylab = "residuals")
abline(0,0)


# QQPlot for group-level effects
qqnorm(ranef(m4)$Bee.ID[[1]], main="Normal Q-Q plot for random effects")
qqline(ranef(m4)$Bee.ID[[1]]) # looks good
```

## Wing Velocity

```{r}
# make a full model with all two-way interactions
m1 <- lmer(wing.velocity ~  (size_pca1 +  Treatment.order + percLoad)^2 + (1|Bee.ID), data = bdta)

summary(m1)


### LRT's for interactions
m2.0 <- update(m1, .~. - Treatment.order:percLoad )

anova(m1, m2.0)
summary(m2.0)


m2.1 <- update(m2.0,.~. - size_pca1:Treatment.order)
anova(m2.0, m2.1)
summary(m2.1)


m2.2 <- update(m2.1, .~. - size_pca1:percLoad)
anova(m2.1, m2.2) ## don't drop size:percLoad interaction


# renaming model to simplify later typing
m2 <- m2.1

##### LRTs for main effects
# trt order
m3 <- update(m2, .~. - Treatment.order)
anova(m2, m3, test = "Chi") # drop trt order
summary(m3)



# summarize final model for paper
summary(m3)

# write output
summary(m3)$coefficients  
write.csv( summary(m3)$coefficients, file = "VelocCoefs_percLoad.csv" )


# diagnostics
# qq plot
qqnorm(resid(m3), main = "")
qqline(resid(m3)) # good

# cook's distance
plot(influence(m3, obs = TRUE), which = 'cook')

# residual plot
plot(fitted(m3), resid(m3), xlab = "fitted", ylab = "residuals")
abline(0,0)


# QQPlot for group-level effects
qqnorm(ranef(m3)$Bee.ID[[1]], main="Normal Q-Q plot for random effects")
qqline(ranef(m3)$Bee.ID[[1]]) # looks good

```

##########################################################################
##
## Part II: Which flight measurements are most closely associated with 
## respiratory rate in bees – wingbeat frequency, stroke amplitude, or 
## wing velocity?
##
##########################################################################

### Multilevel model approach to see the relatedness of the response variables
- Wingbeat frequency
- Stroke amplitude
- Wing velocity
- Resp. Rate



```{r}
# predicting avg resp rate, including stroke amplitude, wing velocity and wingbeat frequency
# looks like we have to remove wing velocity
vif.mer(lmer(av.resp..CO2.mL.hr. ~ stroke.amplitude + wing.velocity + wbf.aud. + percLoad + Treatment.order + size_pca1 + (1|Bee.ID), data = bdta))

vif.mer(lmer(av.resp..CO2.mL.hr. ~ stroke.amplitude + wbf.aud. + percLoad + Treatment.order + size_pca1 + (1|Bee.ID), data = bdta)) # still worryingly high for percLoad and stroke amplitude



mm1 <- lmer(av.resp..CO2.mL.hr. ~ (stroke.amplitude + wbf.aud. + percLoad + Treatment.order + size_pca1)^2 + (1|Bee.ID), data = bdta)
summary(mm1)  # warning says to rescale

# rescale numeric predictors
bdta_scaled = scale(bdta[, c("av.resp..CO2.mL.hr.", 
                             "stroke.amplitude", 
                             "wbf.aud.",
                             "percLoad",
                             "size_pca1")], 
                    center = TRUE, 
                    scale = TRUE)
colnames(bdta_scaled) <- paste0(colnames(bdta_scaled), "_scaled")
bdta = cbind(bdta, bdta_scaled)

mm1 <- lmer(av.resp..CO2.mL.hr. ~ (stroke.amplitude_scaled + wbf.aud._scaled + percLoad + Treatment.order + size_pca1_scaled)^2 + (1|Bee.ID), data = bdta)
summary(mm1)

# LRT's
mm2 <- update(mm1, .~. - wbf.aud._scaled:size_pca1_scaled)
anova(mm1, mm2)
summary(mm2)

mm3 <- update(mm2, .~. -Treatment.order:size_pca1_scaled) # keep interaction
anova(mm3, mm2)
summary(mm3)

mm4 <- update(mm2, .~. - wbf.aud._scaled:Treatment.order)
anova(mm2, mm4)
summary(mm4)

mm5 <- update(mm4, .~. - stroke.amplitude_scaled:wbf.aud._scaled)
anova(mm4, mm5)
summary(mm5)

mm6 <- update(mm5, .~. - stroke.amplitude_scaled:Treatment.order)
anova(mm5, mm6)
summary(mm6)

mm7 <- update(mm6, .~. - percLoad:Treatment.order)
anova(mm6, mm7)
summary(mm7)

mm8 <- update(mm7, .~. - wbf.aud._scaled:percLoad)
anova(mm8, mm7)
summary(mm8)

mm9 <- update(mm8, .~. - percLoad:size_pca1_scaled)
anova(mm9, mm8)
summary(mm9)

mm10 <- update(mm9, .~. - stroke.amplitude_scaled:percLoad)
anova(mm10, mm9) # keep interaction
summary(mm10)

mm11 <- update(mm9, .~. - stroke.amplitude_scaled:size_pca1_scaled)
anova(mm9, mm11) # drop interaction
summary(mm11)


## main effects
mm14 <- update(mm11, .~. - wbf.aud._scaled)
anova(mm11, mm14) # wbf is significant
summary(mm11)


summary(mm11)  # final model for paper


# write output
summary(mm11)$coefficients  
write.csv( summary(mm11)$coefficients, file = "resp_oth_Coefs_percLoad.csv" )

# diagnostics
# qq plot
qqnorm(resid(mm11), main = "")
qqline(resid(mm11)) # good, but two outliers

# residual plot
plot(fitted(mm11), resid(mm11), xlab = "fitted", ylab = "residuals")
abline(0,0)


# QQPlot for group-level effects
qqnorm(ranef(mm11)$Bee.ID[[1]], main="Normal Q-Q plot for random effects")
qqline(ranef(mm11)$Bee.ID[[1]]) # looks good
```

This model is fundamentally different from the previous one.
It's saying, when we account for differences in wingbeat freq, stroke amplitude, and bee size, we sill see an effect of percent load on metabolic rate.

OR, we could say that holding all other variables constant, an increase in wingbeat frequency is associated with an increase in metabolic rate. 


### REFREF: Analyze for load (or percent load)

### Analyze total mass is associated with metabolic rate ? (REFREF)


