---
title: "Bee Respiration for Susie and Stacey"
author: "Callin Switzer"
date: "11/4/2016, Update: 12/13/2016"
output: html_document
---

Analyses of respirometry data:

1.	How does carrying a load during flight affect bumblebees’?
     a.	How does load affect respiratory rate?
     b.	How does load affect wingbeat frequency?
     c.	How does load affect stroke amplitude?
     d.	How does load affect wing velocity?

2.	Which flight measurements are most closely associated with respiratory rate in bees – wingbeat frequency, stroke amplitude, or wing velocity?

## TODO: 
Make better notes
double check visualizations
compare  effect size
Compare sizepca^2


---

Install required packages and read in data
Define custom function for evaluating VIF with multilevel models

```{r}
ipak <- function(pkg){
     new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
     if(length(new.pkg)) install.packages(new.pkg, dependencies = TRUE)
     sapply(pkg, require, character.only = TRUE)
}

packages <- c('ggplot2', 'car', 'lme4', 'gsheet', "MASS", 
              'influence.ME', 'sjPlot', "effects", 'visreg')
ipak(packages)

# read in data -- google sheet called "Bumble mumble grumble"
bdta <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1GUUoFq41Ep3sJNFXiyMcp7fZhYmQCzTcLnhVXdr4WRo/edit?usp=sharing")
```

```{r}

vif.mer <- function (fit) {
    ## adapted from rms::vif
    v <- vcov(fit)
    nam <- names(fixef(fit))

    ## exclude intercepts
    ns <- sum(1 * (nam == "Intercept" | nam == "(Intercept)"))
    if (ns > 0) {
        v <- v[-(1:ns), -(1:ns), drop = FALSE]
        nam <- nam[-(1:ns)]
    }
    
    d <- diag(v)^0.5
    v <- diag(solve(v/(d %o% d)))
    names(v) <- nam
    v
}

```


## Data Overview
## Data cleaning and calculations
```{R}
# create centered and squared variables
bdta <- within(bdta, {
     # new variables
     percLoad <- (Mstarved + load) / Mstarved
     totalMass <- Mstarved + load})

bdta <- within(bdta, {
     # centered variables
     percLoad_cent <- scale(bdta$percLoad, center = TRUE, scale = FALSE)[,1]
     totalMass_cent <- scale(totalMass, center = TRUE, scale = FALSE)[,1]
     wbf_cent <- scale(wbf.aud., center = TRUE, scale = FALSE)[,1]
     strokeAmp_cent <- scale(stroke.amplitude, center = TRUE, scale = FALSE)[,1]
     wingVel_cent <- scale(wing.velocity, center = TRUE, scale = FALSE)[,1]
     
     #squared
     percLoad2 <- percLoad_cent^2
     totalMass2 <- totalMass^2
     wbf2 <- wbf_cent^2  
     strokeAmp2 <- strokeAmp_cent^2
     wingVel2 <- wingVel_cent^2
     
     # change factor properties
     # convert trt order to factor
     Treatment.order <- as.factor(as.character(Treatment.order))

     # change reference level to unloaded
     Treatment <- factor(Treatment, levels = c("UL", "L"))
})

summary(bdta)

ggplot(bdta, aes(x = percLoad, fill = Treatment)) + 
     geom_histogram()  # seems like percent load and trt are very related

ggplot(bdta, aes(x = totalMass, fill = Treatment)) + 
     geom_histogram()  

```

```{r}
# see number of observations per bee
table(bdta$Bee.ID)

# get number of unque bees
length(table(bdta$Bee.ID))

# get treatment orders
trtOrders_loadedSecond <- sapply(X = unique(bdta$Bee.ID), FUN = function(x){
     tmp = bdta[bdta$Bee.ID == x, c("Treatment.order", "Treatment")]
     if("2_L" %in% paste(tmp[,1], tmp[,2], sep = "_")){
          loadedSecond = TRUE
     }
     else loadedSecond = FALSE
     return(loadedSecond)
})

table(trtOrders_loadedSecond)



# visualize scatterplot
car::scatterplotMatrix(bdta[, 4:11])
```

---
Part I: How does carrying a load during flight affect bumblebees’?
a.	How does load affect respiratory rate?
b.	How does load affect wingbeat frequency?
c.	How does load affect stroke amplitude?
d.	How does load affect wing velocity?
In the models, we'll account for differences in bee size and the order of the treatment
---



## Variance Inflation Factor and PCA

```{r}
# VIF  is high among the bee size predictors
# note: this function is almost the same as car::vif
vif.mer(lmer(av.resp..CO2.mL.hr. ~ Treatment + Treatment.order + Mstarved + IT.Span + single.wing.area + (1|Bee.ID), data = bdta))

car::scatterplotMatrix(bdta[, c("Mstarved", "IT.Span", "single.wing.area")])

# principle components
aa = prcomp(bdta[, c("Mstarved", "IT.Span", "single.wing.area")], center = TRUE, scale = TRUE)
summary(aa) # 1st pc explains ~95% of the variance in the three measurements of size

biplot(aa) # shows that all three size measurement are correlated

# note, I changed the signs of the predictions so that higher PC1 values 
# correspond to bigger bees
p1 = -predict(aa)[,1] 

# add PC1 scores to dataset
bdta$size_pc1 = p1
bdta$size_pc1_2 = bdta$size_pc1^2

# show scatterplot matrix to see correlations among size predictors
car::scatterplotMatrix(bdta[, c("Mstarved", "IT.Span", "single.wing.area", "size_pc1")])

# check VIF one last time
# VIF  is high among the bee size predictors
vif.mer(lmer(av.resp..CO2.mL.hr. ~ Treatment + Treatment.order + size_pc1 + size_pc1_2 + 
                  (1|Bee.ID), data = bdta))
```

## Multilevel models

## Resp Rate

```{r}
# make a full model with all two-way interactions
m1 <- lmer(av.resp..CO2.mL.hr. ~  (size_pc1 +  Treatment.order + Treatment)^2  + size_pc1_2 + 
           + (1|Bee.ID), data = bdta)

summary(m1)


### LRT's for interactions
m2.0 <- update(m1, .~. - Treatment.order:Treatment)

anova(m1, m2.0) # likelihood ratio test for interaction of treatment order
summary(m2.0)


m2.1 <- update(m2.0,.~. - size_pc1:Treatment.order)
anova(m2.0, m2.1)
summary(m2.1)


m2.2 <- update(m2.1, .~. - size_pc1:Treatment)
anova(m2.1, m2.2) ## drop all interactions

summary(m2.2)

m2.3 <- update(m2.2, .~. - size_pc1_2)
anova(m2.1, m2.3) # drop sqared terms

# renaming model to simplify later typing
m2 <- m2.3

##### LRTs for main effects
## Treatment Order
m3 <- update(m2, .~. - Treatment.order)
anova(m2, m3, test = "Chi")

# LRT for size
m4 <- update(m2, .~. - size_pc1)
anova(m2, m4)

# LRT for Treatment (load)
m5 <- update(m2, .~. - Treatment)
anova(m2, m5)

# summarize final model for paper
summary(m2)

# write output
summary(m2)$coefficients  
write.csv( summary(m2)$coefficients, file = "RespCoefs_TrtULVsL.csv" )


# diagnostics
# qq plot
qqnorm(resid(m2), main = "")
qqline(resid(m2)) # good

# residual plot
plot(fitted(m2), resid(m2), xlab = "fitted", ylab = "residuals")
abline(0,0)


# QQPlot for group-level effects
qqnorm(ranef(m2)$Bee.ID[[1]], main="Normal Q-Q plot for random effects")
qqline(ranef(m2)$Bee.ID[[1]]) # looks good

infl <- influence(m2, obs = TRUE)
plot(infl, which = 'cook')
```


### Resp Rate Visualization
```{r}
summary(m2)


# visualize with package, "effects"
plot(effect(term = 'Treatment', mod = m2))

# visualize parital residuals, with package "visreg"
aa <- visreg(m2, xvar = 'Treatment', ylab = 'av.resp..CO2.mL.hr.')
bb <- visreg(m2, xvar = 'Treatment', ylab = 'av.resp..CO2.mL.hr.', cond = list(Treatment.order = 2))

# slightly different from plotting raw data
ggplot(bdta, aes(x = Treatment, y = av.resp..CO2.mL.hr.)) + 
     geom_point(position = position_jitter(width = 0.2))  + 
    geom_boxplot(fill = 'NA')

# plot predictions
newdf <- data.frame(size_pc1 = 0.01058963, Treatment.order = factor(1), Treatment = as.factor(rep(c("L", "UL"), each = 30)), Bee.ID = bdta$Bee.ID)

newdf$Treatment <- relevel(newdf$Treatment, ref = "UL")

newdf$preds1 <- predict(m2, newdf)

# plot of prediction for average sized bee (line)
# raw data plotted as points
ggplot(newdf, aes(x = Treatment , y = preds1)) + 
     geom_point(alpha = 0.3, position = position_jitter(width = 0.2))


```


## Wingbeat Freq

```{r}
# make a full model with all two-way interactions
m1 <- lmer(wbf.aud. ~  (size_pc1 +  Treatment.order + Treatment)^2 + (1|Bee.ID), data = bdta)

summary(m1)


### LRT's for interactions
m2.0 <- update(m1, .~. - size_pc1:Treatment )

anova(m1, m2.0)
summary(m2.0)


m2.1 <- update(m2.0,.~. - size_pc1:Treatment.order)
anova(m2.0, m2.1)
summary(m2.1)


m2.2 <- update(m2.1, .~. - Treatment.order:Treatment)
anova(m2.1, m2.2) ## drop all interactions

summary(m2.2)


# renaming model to simplify later typing
m2 <- m2.2

##### LRTs for main effects
## Treatment Order
m3 <- update(m2, .~. - Treatment.order)
anova(m2, m3, test = "Chi")

# LRT for size
m4 <- update(m2, .~. - size_pc1)
anova(m2, m4)

# LRT for Treatment (load)
m5 <- update(m2, .~. - Treatment)
anova(m2, m5)

# summarize final model for paper
summary(m2)

# write output
summary(m2)$coefficients  
write.csv( summary(m2)$coefficients, file = "FreqCoefs_TrtULVsL.csv" )


# diagnostics
# qq plot
qqnorm(resid(m2), main = "")
qqline(resid(m2)) # good

# residual plot
plot(fitted(m2), resid(m2), xlab = "fitted", ylab = "residuals")
abline(0,0)


# QQPlot for group-level effects
qqnorm(ranef(m2)$Bee.ID[[1]], main="Normal Q-Q plot for random effects")
qqline(ranef(m2)$Bee.ID[[1]]) # looks good

```



## Stroke Amplitude
```{r}
# make a full model with all two-way interactions
m1 <- lmer(stroke.amplitude ~  (size_pc1 +  Treatment.order + Treatment)^2 + (1|Bee.ID), data = bdta)

summary(m1)


### LRT's for interactions
m2.0 <- update(m1, .~. - Treatment.order:Treatment )

anova(m1, m2.0)
summary(m2.0)


m2.1 <- update(m2.0,.~. - size_pc1:Treatment.order)
anova(m2.0, m2.1)
summary(m2.1)


m2.2 <- update(m2.1, .~. - size_pc1:Treatment)
anova(m2.1, m2.2) ## drop all interactions

summary(m2.2)


# renaming model to simplify later typing
m2 <- m2.2

##### LRTs for main effects
## size
m3 <- update(m2, .~. - size_pc1)
anova(m2, m3, test = "Chi") # drop size_pc1
summary(m3)

# trt order
m4 <- update(m3, .~. - Treatment.order)
anova(m3, m4)

# LRT for Treatment (load)
m5 <- update(m3, .~. - Treatment)
anova(m3, m5)

# summarize final model for paper
summary(m3)

# write output
summary(m3)$coefficients  
write.csv( summary(m3)$coefficients, file = "AmpCoefs_TrtULVsL.csv" )


# diagnostics
# qq plot
qqnorm(resid(m3), main = "")
qqline(resid(m3)) # ok

# residual plot
plot(fitted(m3), resid(m3), xlab = "fitted", ylab = "residuals")
abline(0,0)


# QQPlot for group-level effects
qqnorm(ranef(m3)$Bee.ID[[1]], main="Normal Q-Q plot for random effects")
qqline(ranef(m3)$Bee.ID[[1]]) # looks good
```

## Wing Velocity

```{r}
# make a full model with all two-way interactions
m1 <- lmer(wing.velocity ~  (size_pc1 +  Treatment.order + Treatment)^2 + (1|Bee.ID), data = bdta)

summary(m1)


### LRT's for interactions
m2.0 <- update(m1, .~. - Treatment.order:Treatment )

anova(m1, m2.0)
summary(m2.0)


m2.1 <- update(m2.0,.~. - size_pc1:Treatment.order)
anova(m2.0, m2.1)
summary(m2.1)


m2.2 <- update(m2.1, .~. - size_pc1:Treatment)
anova(m2.1, m2.2) ## don't drop size:trt interaction


# renaming model to simplify later typing
m2 <- m2.1

##### LRTs for main effects
# trt order
m3 <- update(m2, .~. - Treatment.order)
anova(m2, m3, test = "Chi") # drop trt order
summary(m3)


# summarize final model for paper
summary(m3)

# write output
summary(m3)$coefficients  
write.csv( summary(m3)$coefficients, file = "VelocCoefs_TrtULVsL.csv" )


# diagnostics
# qq plot
qqnorm(resid(m3), main = "")
qqline(resid(m3)) # good

# residual plot
plot(fitted(m3), resid(m3), xlab = "fitted", ylab = "residuals")
abline(0,0)


# QQPlot for group-level effects
qqnorm(ranef(m3)$Bee.ID[[1]], main="Normal Q-Q plot for random effects")
qqline(ranef(m3)$Bee.ID[[1]]) # looks good

```

##########################################################################
##
## Part II: Which flight measurements are most closely associated with 
## respiratory rate in bees – wingbeat frequency, stroke amplitude, or 
## wing velocity?
##
##########################################################################

### Multilevel model approach to see the relatedness of the response variables
- Wingbeat frequency
- Stroke amplitude
- Wing velocity
- Resp. Rate



```{r}
# predicting avg resp rate, including stroke amplitude, wing velocity and wingbeat frequency
# looks like we have to remove wing velocity
vif.mer(lmer(av.resp..CO2.mL.hr. ~ stroke.amplitude + wing.velocity + wbf.aud. + Treatment + Treatment.order + size_pc1 + (1|Bee.ID), data = bdta))

vif.mer(lmer(av.resp..CO2.mL.hr. ~ stroke.amplitude + wbf.aud. + Treatment + Treatment.order + size_pc1 + (1|Bee.ID), data = bdta)) # still worryingly high for treatment and stroke amplitude



mm1 <- lmer(av.resp..CO2.mL.hr. ~ (stroke.amplitude + wbf.aud. + Treatment + Treatment.order + size_pc1)^2 + (1|Bee.ID), data = bdta)
summary(mm1)  # warning says to rescale

# rescale numeric predictors
bdta_scaled = scale(bdta[, c("av.resp..CO2.mL.hr.", 
                             "stroke.amplitude", 
                             "wbf.aud.", 
                             "size_pc1")], 
                    center = TRUE, 
                    scale = TRUE)
colnames(bdta_scaled) <- paste0(colnames(bdta_scaled), "_scaled")
bdta = cbind(bdta, bdta_scaled)

mm1 <- lmer(av.resp..CO2.mL.hr. ~ (stroke.amplitude_scaled + wbf.aud._scaled + Treatment + Treatment.order + size_pc1_scaled)^2 + (1|Bee.ID), data = bdta)
summary(mm1)

# LRT's
mm2 <- update(mm1, .~. - wbf.aud._scaled:size_pc1_scaled)
anova(mm1, mm2)
summary(mm2)

mm3 <- update(mm2, .~. -Treatment.order:size_pc1_scaled)
anova(mm3, mm2)
summary(mm3)

mm4 <- update(mm3, .~. - wbf.aud._scaled:Treatment.order)
anova(mm3, mm4)
summary(mm4)

mm5 <- update(mm4, .~. - stroke.amplitude_scaled:wbf.aud._scaled)
anova(mm4, mm5)
summary(mm5)

mm6 <- update(mm5, .~. - stroke.amplitude_scaled:Treatment.order)
anova(mm5, mm6)
summary(mm6)

mm7 <- update(mm6, .~. - Treatment:Treatment.order)
anova(mm6, mm7)
summary(mm7)

mm8 <- update(mm7, .~. - wbf.aud._scaled:Treatment)
anova(mm8, mm7)
summary(mm8)

mm9 <- update(mm8, .~. - Treatment:size_pc1_scaled)
anova(mm9, mm8)
summary(mm9)

mm10 <- update(mm9, .~. - stroke.amplitude_scaled:Treatment)
anova(mm10, mm9)
summary(mm10)

mm11 <- update(mm10, .~. - stroke.amplitude_scaled:size_pc1_scaled)
anova(mm10, mm11) # keep interaction

## main effects
mm12 <- update(mm10, .~. - Treatment.order)
anova(mm10, mm12)
summary(mm12)

## test only main effects that aren't part of the interaction
mm13 <- update(mm12, .~. -Treatment)
anova(mm12, mm13) # trt is significant


mm14 <- update(mm12, .~. - wbf.aud._scaled)
anova(mm12, mm14) # wbf is significant

mm15 <- update(mm12, .~. - stroke.amplitude_scaled:size_pc1_scaled)
anova(mm12, mm15)


summary(mm12)  # final model for paper


# write output
summary(mm12)$coefficients  
write.csv( summary(mm12)$coefficients, file = "resp_oth_Coefs_TrtULVsL.csv" )

# diagnostics
# qq plot
qqnorm(resid(mm12), main = "")
qqline(resid(mm12)) # good, but two outliers

# residual plot
plot(fitted(mm12), resid(mm12), xlab = "fitted", ylab = "residuals")
abline(0,0)


# QQPlot for group-level effects
qqnorm(ranef(mm12)$Bee.ID[[1]], main="Normal Q-Q plot for random effects")
qqline(ranef(mm12)$Bee.ID[[1]]) # looks good
```

This model is fundamentally different from the previous one.
It's saying, when we account for differences in wingbeat freq, stroke amplitude, and bee size, we sill see an effect of treatment on metabolic rate.

OR, we could say that holding all other variables constant, an increase in wingbeat frequency is associated with an increase in metabolic rate. 

